Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2016-12-22 manuscript.tex
latest

\begin{figure}
\centering
\includegraphics[scale=0.8]{lottery.pdf}
\caption{\label{fig:lottery} Each iteration of our lottery model has three main elements. First, propagules are produced by adults which are dispersed at random over the unoccupied territories (only propagules landing on unoccupied territories are shown). Lottery competition then occurs in each unoccupied territory: each genotype has a probability proportional to $b_i n_i c_i$ of securing the territory. Then occupied territories are freed up by adult mortality. In Eq. \eqref{eq:delttot} and most of the paper, only adults can die (red crosses), but we will also consider the case where juveniles die (blue cross; section ``Primary strategies and Grime's triangle'').}
\end{figure}

@@

The stress archetype is more ambiguous, and has been the subject of an extensive debate in the plant ecology literature (the ``Grime-Tilman'' debate; \citealt{aerts_1999} and references therein). Stressful environments severely restrict growth and reproduction, so that $b_i\ll 1$ \cite{grime_1974,grime_1977}. Mutations which appreciably improve $b_i$ will be either non-existent or extremely unlikely, so $b_i$ is constrained to remain low. In Grime's view, under these conditions the rate at which propagules successfully develop to adulthood cannot appreciably exceed the mortality rate. This implies $b_i/d_i\approx 1$ in our model, and so the propagule density $L$ is suppressed to such low levels that there are essentially no territorial contests occurring.

The alternative view is that stressful environments simply have a lower carrying capacity \citep{taylor_1990}; in our model, this means a greater per-individual territorial requirement represented by a lower $T$ for a given amount of space. In particular, it is argued that when stress is induced by a scarcity of consumable resources, competition for those resources would likely be intense (for empirical support, see \citealt{davis_1998}). Thus, $b_i$ need not be particularly close to $d_i$, but contests for consumable resources at the juvenile phase would kill off most propagules before adulthood. In other words, the stressed population is a high density population where competition is important (that is, high density relative to $T$, not relative to ideal conditions; \citealt{taylor_1990}).

>>>>

\begin{figure}
\centering
\includegraphics[scale=0.8]{lottery.pdf}
\caption{\label{fig:lottery} Each iteration of our lottery model has three main elements. First, propagules are produced by adults which are dispersed at random over the unoccupied territories (only propagules landing on unoccupied territories are shown). Lottery competition then occurs in each unoccupied territory (only competition in one territory is illustrated): each genotype has a probability proportional to $b_i n_i c_i$ of securing the territory. Then occupied territories are freed up by adult mortality. In Eq. \eqref{eq:delttot} and most of the paper, only adults can die (red crosses), but we will also consider the case where juveniles die (blue cross; section ``Primary strategies and Grime's triangle'').}
\end{figure}

@@

The stress archetype is more ambiguous, and has been the subject of an extensive debate in the plant ecology literature (the ``Grime-Tilman'' debate; \citealt{aerts_1999} and references therein). Stressful environments severely restrict growth and reproduction, so that $b\ll 1$ \cite{grime_1974,grime_1977}. Mutations which appreciably improve $b$ will be either non-existent or extremely unlikely, so $b$ is constrained to remain low. In Grime's view, under these conditions the rate at which propagules successfully develop to adulthood cannot appreciably exceed the mortality rate. This implies $b/d\approx 1$ in our model, and so the propagule density $L$ is suppressed to such low levels that there are essentially no territorial contests occurring.

The alternative view is that, while stressful environments imply lower $b$ and support a lower number of individuals per unit area compared what is attainable in ideal environments, stressed populations are actually at high densities relative to the environmental carrying capacity, and are highly competitive \citep{taylor_1990}. In the particular case that stress is caused by scarcity of consumable resources, it is argued that we should expect intense resource competition (for empirical support, see \citealt{davis_1998}). Thus, $b$ may actually appreciably exceed $d$ under stressful conditions, even though the absolute value of $b$ is small.
